---
title: "Assembly diet based minimal microbiome (Db-MM10)"
subtitle: "RNASeq processing"
author: "Sudarshan A. Shetty"
date: "`r date()`"
output:
  workflowr::wflow_html:
    toc: yes
    toc_depth: 2
    toc_float: true
editor_options: 
  chunk_output_type: console
---


# Introduction  

The analysis here is based on triplicates.    
Read outputs of DIAMOND analysis into R. The code follows the R scripts from SAMSA2 pipeline.  

## Setup

```{r message=FALSE, warning=FALSE}

# import other necessary packages
#library(optparse)
library(tidyverse)
#library(EnhancedVolcano)
library(ggplot2)


```


# Read counts  

Input files here were obtained from SAMSA2 pipeline.
## Get file names  
```{r}


# GET FILE NAMES Control 24h
control_files <- list.files(path = "data_raw/04_metatrans/samsa_output",
                            pattern = "control_*", 
                            full.names = T, recursive = FALSE)

control_names <- ""
for (name in control_files) {
  
  x <- gsub("data_raw/04_metatrans/samsa_output/", "", name)
  x <- gsub(".merged.RefSeq_annotated_BLAST.txt", "", x)
  control_names <- c(control_names, x)
}
control_names

control_names_trimmed <- control_names[-1]

# Experimental 48h

exp_files <- list.files(path = "data_raw/04_metatrans/samsa_output",
                        pattern = "experimental_*",
                        full.names = T,
                        recursive = FALSE)
exp_names = ""
for (name in exp_files) {
  
  x <- gsub("data_raw/04_metatrans/samsa_output/", "", name)
  x <- gsub(".merged.RefSeq_annotated_BLAST.txt", "", x)
  exp_names <- c(exp_names, x)
}


exp_names_trimmed <- exp_names[-1]

# sanity check
if (length(exp_files) == 0 || length(control_files) == 0) {
  print ("\nWARNING: No files found.  Is the directory correct?  Are the files named with 'control_' and 'experimental_' as prefixes?")
  stop()
}

#head(control_names_trimmed)
#head(exp_names_trimmed)

```

## Read files  

```{r}

# loading the control table
y <- 0
for (x in control_files) {
  y <- y + 1
  if (y == 1) {
    control_table <- read.table(file = x, header = F, quote = "", sep = "\t")
    colnames(control_table) = c(x, "V2")
    control_table <- control_table[,c(2,1)]      }
  if (y > 1) {
    temp_table <- read.table(file = x, header = F, quote = "", sep = "\t")
    colnames(temp_table) = c( x, "V2")
    temp_table <- temp_table[,c(1,2)]
    control_table <- merge(control_table, temp_table, by = "V2", all = T)  }
}

#head(control_table)

control_table[is.na(control_table)] <- 0
rownames(control_table)= control_table$V2
 
control_table_trimmed <- data.frame(control_table[,-1])

# loading the experimental table
z <- 0
for (x in exp_files) {
  z <- z + 1
  if (z == 1) {
    exp_table <- read.table(file = x, header=F, quote = "", sep = "\t")
    colnames(exp_table) = c(x, "V2")
    exp_table <- exp_table[,c(2,1)]  }
  if (z > 1) {
    temp_table <- read.table(file = x, header = F, quote = "", sep = "\t")
    colnames(temp_table) = c(x, "V2")
    exp_table <- merge(exp_table, temp_table[,c(1,2)], by = "V2", all = T)  }
}
exp_table[is.na(exp_table)] <- 0
rownames(exp_table) = exp_table$V2
exp_table_trimmed <- exp_table[,-1]

```

check control_table 24h
```{r}
head(control_table_trimmed)
```


check control_table_trimmed 48h
```{r}
head(exp_table_trimmed)
```



Format the two tables.   
## Create main table  
```{r}

# getting the column names simplified
colnames(control_table_trimmed) = control_names_trimmed
colnames(exp_table_trimmed) = exp_names_trimmed

# merging the two tables (time-consuming step)
complete_table <- merge(control_table_trimmed, exp_table_trimmed, by=0, all = TRUE)
complete_table[is.na(complete_table)] <- 0
# reducing stuff down to avoid duplicates
complete_table <- aggregate(. ~  Row.names, data = complete_table, sum)

rownames(complete_table) <- complete_table$Row.names

complete_table <- complete_table[!(complete_table$Row.names == ""), ]

# removing extra Row.names column
complete_table <- complete_table[,-1]
complete_table.a <- complete_table
#colnames(complete_table.a)
#head(complete_table)

colnames(complete_table) <- gsub("control_", "", colnames(complete_table))
colnames(complete_table) <- gsub("experimental_", "", colnames(complete_table))
```


Store the complete table for later analysis.  

## Save objects  
```{r}

raw_counts_table <- complete_table
saveRDS(raw_counts_table, "data/04_metatrans/rds/raw_counts_tables.rds")

write.table(raw_counts_table, "data/04_metatrans/tables/raw_counts_tables.txt", col.names=NA, sep = "\t")

```

Here, the `raw_counts_tables.rds` contains locus tag counts in each sample.  

```{r eval=FALSE}
sessionInfo()
```

