---
title: "Assembly diet based minimal microbiome (Db-MM10)"
subtitle: "Culture experimemnts and Metabolites anlaysis"
author: "Sudarshan A. Shetty"
date: "`r date()`"
output:
  workflowr::wflow_html:
    toc: yes
    toc_depth: 2
    toc_float: true
editor_options: 
  chunk_output_type: console
---


# Introduction  
This document consist of codes for analysis and visualization of lab data from growth assays, hplc measurement of individual and mixed cultures of Db-MM. 

## Set up
```{r, message=FALSE}
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(microbiome)
library(patchwork)
source('code/custom_funs.R')

```

Define a color for metabolites tested in the Db-MM  

```{r}
# for metabolites tested in the Db-MM  
ColorsDT <- c(Acetate = "#CBD588", Propionate = "#5F7FC7", 
              Butyrate = "orange", Isobutyrate= "#DA5724", 
              Formate ="#508578", Lactate = "#CD9BCD", Glucose = "#AD6F3B", 
              Fructose = "#673770",Xylose = "#D14285", Cellobiose= "black", 
              Hydrogen_gas="#a8ddb5", Hydrogen = "#08306b")

```


# Mono Culture  

```{r}

hplc_data <- read.csv("data_raw/02_wetlab/mono_culture_hplc.csv", 
                      row.names = 1, header = T, 
                      stringsAsFactors = F)

hplc_data$BacterialStrain <- gsub("Lachnospiraceae bacterium 7_1_58FAA", 
                                  "Flavonifractor plautii",
                                  hplc_data$BacterialStrain)
head(hplc_data)
```

```{r}
all2 <- reshape2::melt(hplc_data)
all2 <- subset(all2, variable != "Isobutyrate") # below detection limit
```


## Acetate concentrations  
```{r}
## Since acetate was added in the media and had high concentration, I will plot it separately 
all2.acetate <- subset(all2, variable == "Acetate")

all2.acetate$TimepointNum <- all2.acetate$Timepoint
all2.acetate$TimepointNum <- as.numeric(gsub("T", "",all2.acetate$TimepointNum))

p.ace <- ggline(all2.acetate, 
                x ="TimepointNum", 
                y = "value", 
                color = "BacterialStrain", 
                add = "mean_se",
                ylab = "Acetate concentration (mM)",
                xlab = "Time (hr)", 
                legend = "right", 
                size = 0.7,
                alpah= 0.5,
                ncol =5,
                add.params = list(alpha = 0.75)) + 
  rotate_x_text() + 
  scale_color_manual("", values= strain.colors) +
  theme_minimal(base_size = 10) + 
  theme(#strip.background = element_rect(fill="white"),
    strip.background = element_rect(color= "grey10"), 
    strip.text = element_text(face = "italic"),
    panel.border = element_rect(colour = "grey10", fill=NA),
    axis.text.x=element_text(colour="grey10", size = 11),
    axis.text.y=element_text(colour="grey10", size = 11),
    axis.title = element_text(colour="grey10", size = 11),
    panel.spacing = unit(0.1, "lines")) 
p.ace <- ggpar(p.ace, legend = "right", font.legend = c(8, "italic"))
p.ace
ggsave("data/02_wetlab/figs/R1_mono_culture_HPLC_Acetate.pdf", height = 4, width = 8)


```


## Remaining metabolites  
```{r}

all2.noace <- subset(all2, variable !="Acetate")

all2.noace$TimepointNum <- all2.noace$Timepoint
all2.noace$TimepointNum <- as.numeric(gsub("T", "",all2.noace$TimepointNum))
#head(hplc_metadata)

p1 <- ggline(all2.noace, 
             x = "TimepointNum", 
             y = "value", 
             color = "variable", 
             add = "mean_se",
             facet.by = "BacterialStrain",
             ylab = "Concentration (mM)",
             xlab = "Time (hr)", 
             legend = "right", #scales ="free",
             ncol = 5,
             size = 0.7,
             alpha=0.5) + rotate_x_text() + 
  scale_color_manual("", values= ColorsDT) +
  theme_minimal(base_size = 10) + 
  theme(#strip.background = element_rect(fill="white"),
    strip.background = element_rect(color= "grey10"), 
    strip.text = element_text(face = "italic"),
    panel.border = element_rect(colour = "grey10", fill=NA),
    axis.text.x=element_text(colour="grey10", size = 11),
    axis.text.y=element_text(colour="grey10", size = 11),
    axis.title = element_text(colour="grey10", size = 11),
    panel.spacing = unit(0.1, "lines")) +  
  geom_hline(aes(yintercept=4), linetype="dashed", alpha = 0.5) 
p1 


ggsave("data/02_wetlab/figs/R1_mono_culture_HPLC_metabolites.pdf", height = 5, width = 14)
#ggsave("output/figs/01_mono_culture_HPLC_metabolites.png", height = 5, width = 14)
#ggsave("output/figs/01_mono_culture_HPLC_metabolites.tiff", height = 5, width = 14)

```

# Reviewer-2  
Split sugars and SCFAs  

## Sugars  
```{r}
#Butyrate   Formate    Lactate    Propionate Cellobiose Fructose   Glucose    Xylose
sug <- c("Fructose", "Glucose", "Xylose")
sugs.all <- all2.noace %>% 
  filter(variable %in% sug) %>% 
  ggline(x = "TimepointNum", 
         y = "value", 
         color = "variable", 
         add = "mean_se",
         facet.by = "BacterialStrain",
         ylab = "Concentration (mM)",
         xlab = "Time (hr)", 
         legend = "right", #scales ="free",
         ncol = 5,
         size = 0.75,
         alpha=0.5) + rotate_x_text() + 
  scale_color_manual("", values= ColorsDT[7:9]) +
  theme_minimal(base_size = 10) + 
  theme(#strip.background = element_rect(fill="white"),
    strip.background = element_rect(color= "grey10"), 
    strip.text = element_text(face = "italic"),
    panel.border = element_rect(colour = "grey10", fill=NA),
    axis.text.x=element_text(colour="grey10", size = 11),
    axis.text.y=element_text(colour="grey10", size = 11),
    axis.title = element_text(colour="grey10", size = 11),
    panel.spacing = unit(0.1, "lines")) 
sugs.all
ggsave("data/02_wetlab/figs/R1_mono_culture_HPLC_sugars.pdf", height = 4, width = 12)
```

## VFAs    
```{r}
scfa.all <- all2.noace %>% 
  filter(!variable %in% c(sug, "Cellobiose")) %>% 
  ggline( "TimepointNum", "value", color = "variable", 
          add = "mean_se",
          facet.by = "BacterialStrain",
          ylab = "Concentration (mM)",
          xlab = "Time (hr)", 
          legend = "right", #scales ="free",
          ncol = 5,
          size = 0.75,
          alpha=0.5) + rotate_x_text() + 
  scale_color_manual("", values= ColorsDT[c(1:3,5,6)]) +
  theme_minimal()+
  theme(#strip.background = element_rect(fill="white"),
    strip.background = element_rect(color= "grey10"), 
    strip.text = element_text(face = "italic"),
    panel.border = element_rect(colour = "grey10", fill=NA),
    axis.text.x=element_text(colour="grey10", size = 11),
    axis.text.y=element_text(colour="grey10", size = 11),
    axis.title = element_text(colour="grey10", size = 11),
    panel.spacing = unit(0.1, "lines"))  
  #geom_hline(aes(yintercept=4), linetype="dashed", alpha = 0.5) 
scfa.all
ggsave("data/02_wetlab/figs/R1_mono_culture_HPLC_SCFA.pdf", height = 4, width = 12)
```

```{r}
#p.ace.n + p.ace.leg + plot_layout(heights = c(2,1))

scfa.all / (p.ace + plot_spacer()) + plot_layout(heights = c(2,1), guides = "collect")
ggsave("data/02_wetlab/figs/R1_mono_culture_SCFA_Acetate.pdf", height = 6, width = 12)
```

# Mixed culture  

```{r}

mixed_data <- read.csv("data_raw/02_wetlab/mixed_culture_scfa_gc.csv", 
                       row.names = 1, header = T, 
                       stringsAsFactors = F)

mixed_data_2 <- reshape2::melt(mixed_data)

mixed_data_2$TimepointNum <- mixed_data_2$Timepoint
mixed_data_2$TimepointNum <- as.numeric(gsub("T", "",mixed_data_2$TimepointNum))

p4 <- ggline(mixed_data_2, "TimepointNum", "value", color = "variable", 
             add = "mean_se", 
             ylab = "Concentration (mM)",
             xlab = "Time (hr)", legend = "right",
             size = 1.5) + 
  rotate_x_text() + 
  scale_color_manual("",values= ColorsDT[c(1,2,3,4,5,6,12)])
p4 <- p4 + geom_vline(aes(xintercept=1.5), linetype="dashed", alpha = 0.5) + geom_vline(aes(xintercept=2.5), linetype="dashed", alpha = 0.7) 
p4 <- p4 + theme_bw(base_size = 14) 
p4

ggsave("data/02_wetlab/figs/R1_Figure_Mixed_HPLC.pdf", height = 3, width = 6)

```

## Reveiwer-1  

```{r}
acetate_t0 <- subset(mixed_data_2, Timepoint == "T0" & variable == "Acetate") %>% 
  summarise(mean.conc = mean(value),
            sd.conc = sd(value))
acetate_t0
#  mean.conc   sd.conc
#  27.87145 0.5910155
```

## Reviewer 2
```{r}
scfa_vals <- subset(mixed_data_2) %>% 
  group_by(variable, Timepoint) %>% 
  summarise(mean.conc = mean(value),
            sd.conc = sd(value))
scfa_vals

readr::write_csv(scfa_vals, "data/02_wetlab/tables/mixed_cultre_scfas.csv")
DT::datatable(scfa_vals)
```

# Optical density and pH  
```{r}
od_metadata <- read.csv("data_raw/02_wetlab/mixed_culture_od_ph.csv", 
                        row.names = 1, header = T, 
                        stringsAsFactors = F)

```


```{r}
# convert to long format
all.OD2 <- reshape2::melt(od_metadata)

all.OD2 <- subset(all.OD2, variable !="Replicate")

od.plot.mix <- all.OD2 %>% 
  filter(variable == "OD600") %>% 
  ggline("Timepoint", "value",
         add = "mean_se",
         xlab = "Time (hr)",
         ylab = expression("O.D"[600]),
         size = 1, 
         linetype = "dashed") + 
  rotate_x_text() + 
  theme_bw(base_size = 14) + 
  ylim(c(0,2.5)) + 
  theme(strip.background = element_rect(fill="white")) 

ph.plot.mix <- all.OD2 %>% 
  filter(variable == "pH") %>% 
  ggline("Timepoint", "value",
         add = "mean_se",
         xlab = "Time (hr)",
         ylab = "pH",
         size = 1, 
         linetype = "dashed") + 
  rotate_x_text() + 
  theme_bw(base_size = 14) + 
  ylim(c(6,8)) + 
  theme(strip.background = element_rect(fill="white")) 
od.plot.mix + ph.plot.mix

ggsave("data/02_wetlab/figs/R1_mixed_OD_PH_figure.pdf", height = 3, width = 6)

```

## Reviewer
```{r}
ph.means <- all.OD2 %>% 
  filter(variable == "pH") %>% 
  group_by(Timepoint) %>% 
  summarise(mean.conc = mean(value),
            sd.conc = sd(value))
ph.means
readr::write_csv(ph.means, "data/02_wetlab/tables/mixed_cultre_ph.csv")

ph.means$Timepoint
```

## Reviewer-2 

```{r}
rint.but <- all2 %>% 
  filter(BacterialStrain == "Roseburia intestinalis") %>% 
  group_by(Timepoint, variable) %>% 
  summarise(mean.conc = mean(value),
            sd.conc = sd(value)) %>% 
  as.data.frame()
rint.but

bac.scfa <- all2 %>% 
  #filter(BacterialStrain == "Roseburia intestinalis") %>% 
  group_by(BacterialStrain, Timepoint, variable) %>% 
  summarise(mean.conc = mean(value),
            sd.conc = sd(value)) %>% 
  as.data.frame()
DT::datatable(bac.scfa)
readr::write_csv(bac.scfa, "data/02_wetlab/tables/mono_cultre_scfas_all.csv")


```

# pH changes in monocultures 
```{r}
# pH 
ph_data <- read.csv("data_raw/02_wetlab/mono_culture_od_ph.csv", 
                    row.names = 1, header = T, 
                    stringsAsFactors = F)
all.ph2 <- reshape2::melt(ph_data)
all.ph2$TimepointNum <- all.ph2$Timepoint
all.ph2$TimepointNum <- as.numeric(gsub("T", "",all.ph2$TimepointNum))

all.ph2$BacterialStrain <- gsub("Lachnospiraceae bacterium 7_1_58FAA", 
                                "Flavonifractor plautii",
                                all.ph2$BacterialStrain)


od.plot.mono <- all.ph2 %>% 
  filter(variable == "OD") %>% 
  ggline(x = "TimepointNum", 
         y = "value",
         color = "BacterialStrain",
         add = "mean_se",
         xlab = "Time (hr)",
         ylab = expression("O.D"[600]),
         size = 1) + 
  rotate_x_text() + 
  scale_color_manual("", values= strain.colors) +
  theme_bw(base_size = 14) + 
  ylim(c(0,3.5)) + 
  theme(strip.background = element_rect(fill="white")) 
od.plot.mono <- ggpar(od.plot.mono, legend = "right", font.legend = c("italic"))
od.plot.mono

ph.plot.mono <- all.ph2 %>% 
  filter(variable == "PH") %>% 
  ggline(x = "TimepointNum", 
         y = "value",
         color = "BacterialStrain",
         add = "mean_se",
         xlab = "Time (hr)",
         ylab = expression("pH"),
         size = 1) + 
  rotate_x_text() + 
  scale_color_manual("", values= strain.colors) +
  theme_bw(base_size = 14) + 
  #ylim(c(4,8)) + 
  theme(strip.background = element_rect(fill="white"),
        legend.position = "none") 
#ph.plot.mono <- ggpar(ph.plot.mono, legend = "right", font.legend = c("italic"))
ph.plot.mono

(od.plot.mono | ph.plot.mono) + plot_layout(guides = "collect")

ggsave("data/02_wetlab/figs/01_mono_culture_OD_PH_New.pdf", height = 4, width = 10)

```


```{r}
sessionInfo()
```

